---
title: "Assignment 10 - Sentiment Analysis"
author: "Ken Popkin"
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r echo = TRUE, message=FALSE, warning=FALSE}
library(tidytext)
library(textdata)
```

### Introduction to Assignment 10
<I> Citation: </I> In Text Mining with R, [Chapter 2](https://www.tidytextmining.com/sentiment.html) looks at Sentiment Analysis. For the assignment below all the code was copied from this source and rerun.  

<I>Assignment:</I>   
Part 1. Get the base code from Text Mining with R, chapter 2 working as-is in this notebook. <br> <br>
Part 2. Extend the code in two ways: <br>
   a. Work with a different corpus <br>
   b. Incorporate at least one additional sentiment lexicon

### Part 1: Get base code working
```{r}
#afinn
afinn_sentiments <- get_sentiments("afinn")

#bing
bing_sentiments <- get_sentiments("bing")

#nrc
nrc_sentiments <- get_sentiments("nrc")
```

```{r echo = TRUE, message=FALSE, warning=FALSE}
library(janeaustenr)
library(dplyr)
library(stringr)

#Put the text in a tidy format
tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", 
                                                 ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}

#Complete the sentiment analysis
#filter for the joy words, then filter() the data frame with the text from the books for the words from Emma and #use inner_join() to perform the sentiment analysis.
nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyr)

jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(ggplot2)

ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")

```

```{r echo=TRUE, message=FALSE, warning=FALSE}
#Section 2.3 Comparing the three sentiment dictionaries
pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")

pride_prejudice
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")


bing_and_nrc <- bind_rows(pride_prejudice %>% 
                            inner_join(get_sentiments("bing")) %>%
                            mutate(method = "Bing et al."),
                          pride_prejudice %>% 
                            inner_join(get_sentiments("nrc") %>% 
                                         filter(sentiment %in% c("positive", 
                                                                 "negative"))) %>%
                            mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)
```

```{r}
#Positive and negative words in each lexicon
get_sentiments("nrc") %>% 
     filter(sentiment %in% c("positive", 
                             "negative")) %>% 
  count(sentiment)


get_sentiments("bing") %>% 
  count(sentiment)
```

```{r}
#Bing word counts
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()


bing_word_counts
```

```{r}
#Bing word count plots
bing_word_counts %>%
  group_by(sentiment) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(y = "Contribution to sentiment",
       x = NULL) +
  coord_flip()
```

```{r}
#How to exclude words to not include (i.e. "miss" is seen as a negative, but could just be a title "Miss Jones")
custom_stop_words <- bind_rows(tibble(word = c("miss"), 
                                          lexicon = c("custom")), 
                               stop_words)

custom_stop_words
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
#Word Clouds
library(wordcloud)

tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```

```{r}
#2.6 Looking at units beyond just words
PandP_sentences <- tibble(text = prideprejudice) %>% 
  unnest_tokens(sentence, text, token = "sentences")

PandP_sentences$sentence[2]
```

```{r}
#Split into a dataframe by chapter
austen_chapters <- austen_books() %>%
  group_by(book) %>%
  unnest_tokens(chapter, text, token = "regex", 
                pattern = "Chapter|CHAPTER [\\dIVXLC]") %>%
  ungroup()


austen_chapters %>% 
  group_by(book) %>% 
  summarise(chapters = n())
```

```{r}
#For each book, which chapter has the highest number of negative words
bingnegative <- get_sentiments("bing") %>% 
  filter(sentiment == "negative")


wordcounts <- tidy_books %>%
  group_by(book, chapter) %>%
  summarize(words = n())


tidy_books %>%
  semi_join(bingnegative) %>%
  group_by(book, chapter) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c("book", "chapter")) %>%
  mutate(ratio = negativewords/words) %>%
  filter(chapter != 0) %>%
  top_n(1) %>%
  ungroup()
```

### Part 2: Extend the code and work with a different corpus
This section extends the above code by working with the Gutenberg corpus, library(guttenbergr), instead of JaneAusten and incorporating the "squzhet" sentiment lexicon to...

1. obtain a work from the Gutenberg corpus
2. run get_sentiments using syuzhet
3. histogram of the positive and negative sentences count
4. create a word cloud

```{r echo=TRUE, warning=FALSE, message=FALSE}
library(gutenbergr)
library(dplyr)
library(stringr)
library(syuzhet)
```
 
<I> Obtain a Work from the Gutenberg Corpus </I>
```{r echo=TRUE, message=FALSE, warning=FALSE}
#Obtain Plato's, "The Republic" corpus
plato <- gutenberg_works(author == "Plato", title == "The Republic")

therepublicid <- plato$gutenberg_id

republic_corpus <- gutenberg_download(therepublicid)

republic_words <- get_tokens(republic_corpus$text, pattern = "\\W")

cat('Word count of The Republic is', length(republic_words), 'words')
```

<I> Get Sentiment using syuzhet </I>
```{r, messages=FALSE, warning=FALSE}
#Get sentiment using syuzhet 
republic_sentences <- get_sentences(republic_corpus$text)

republic_sentiment <- get_sentiment(republic_sentences, method="syuzhet")
```

```{r}
#Count the number of positive and negative sentences in The Republic
pos_sentences <- length(which(republic_sentiment > 0))
neutral_sentences <- length(which(republic_sentiment == 0))
neg_sentences <- length(which(republic_sentiment < 0))
tot_sentences <- length(republic_sentiment)

#In syuzhet positives get a value > 0 and negatives get a value < 0, and 0 is neutral.
cat('Number of positive, neutral, and negative sentences in The Republic is', pos_sentences, neutral_sentences, 'and', neg_sentences, 'respectively for a total of', tot_sentences, 'sentences')
```

<I> Histogram of the Positive and Negative Sentences Count </I>
```{r}
#Histogram of the positive and negative sentences
republic_sentiment_df <- as.data.frame(republic_sentiment)
republic_sentiment_df <- subset(republic_sentiment_df,republic_sentiment != 0)
ggplot(republic_sentiment_df, aes(x=republic_sentiment)) + geom_histogram(bins = 90)
```

<I> Create a Word Cloud </I> 
```{r}
#Create a word cloud of "The Republic"
republic_words <- republic_corpus %>% unnest_tokens(word,text)

republic_words %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```
