---
title: "Home Values Visualization"
author: "Ken Popkin"
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(reshape2)
library(ggpubr)
library(ggfortify)

require(scales)
```


```{r echo=FALSE}
# load data
data <- read.csv('https://raw.githubusercontent.com/krpopkin/Data606_Final_Project_Repository/master/01Data/%20master.csv')

cases = dim(data)[1]
features = dim(data)[2]
cat('Data for this project is', cases, 'cases and', features, 'features')
```

```{r data}
head(data,1)
```

### Part 1 - Introduction
<I> Can education and wages be used as predictors for home values? </I>    
There are many factors theorized to influence a home’s value and for this project we’ll attempt to answer if two popular topics, education and wages, can be used as predictors for a third popular topic, home values.

Knowing key factors to monitor regarding your home’s value can prove beneficial in multiple areas of modern life.  Areas such as assessing your long term financial outlook, deciding whether to move or not, and perhaps most important, enabling these and similar decisions to not be based solely on emotion.  

### Part 2 - Data
The master data file loaded above is a combination of three files ([education](https://nces.ed.gov/ccd/f33agency.asp), [wages](https://www.bls.gov/cew/downloadable-data-files.htm), and [home values](https://www.zillow.com/research/data/)) that were merged using state/county as the key.  For details about the cleaning, filtering, and merging of this data to create the master file refer to the [project proposal](https://www.rpubs.com/krpopkin/591317).

### Part 3 - Exploratory data analysis
### Relevant summary statistics 
<I> Description of the master data file <\I>
```{r}
describe(master)
```

<I> Education <\I>  
More analysis of education will be done in the project.  Below are histograms of high school graduation rates and four year college graduation rates.
```{r}
#Percent with high school diploma
ggplot(master, aes(x=hs_diploma_percent)) + geom_histogram()
```


```{r}
#Percent with four year degrees
ggplot(master, aes(x=four_year_degree_percent)) + geom_histogram()
```

<I> Avg annual pay <\I>  
Average annual pay is skewed right. This makes sense as all US states are in this dataset.
```{r}
ggplot(master, aes(x=avg_annual_pay)) + geom_histogram()
```

<I> Home values plot <\I>  
Similar to wages, home values in some areas are very high, giving the distribution a right skew
```{r}
ggplot(master, aes(x=avg_home_value)) + geom_histogram()
```


<B> Correlation Heat Maps </B>  
We'll start off with heat maps to highlight the correlations between Education, Wages, Resident Count and Home Values.  Our first heatmap will look at the four education categories of high school dropout, high school diploma, some college, and four year degree.  There should be high correlation between each of these so we'll select the one with highest correlation to home values to move forward with.

```{r education-heatmap}
#Select the data for the education heat map
ehm <- data %>%
  select(hs_dropout = dropout_percent, hs_diploma = hs_diploma_percent, some_college = some_college_percent,  college_degree = four_year_degree_percent, avg_home_value)

#Calculate the correlation values
cormat <- round(cor(ehm),2)

melted_cormat <- melt(cormat)

#Create the Heatmap
library(ggplot2)
ggheatmap <- ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()

reorder_cormat <- function(cormat){
# Use correlation between variables as distance
dd <- as.dist((1-cormat)/2)
hc <- hclust(dd)
cormat <-cormat[hc$order, hc$order]
}

ggheatmap + 
geom_text(aes(Var2, Var1, label = value), color = "black", size = 4)
```

The education heatmap correlation of 69% with a college degree and home values was somewhat expected.  The -61% corelation for a high school diploma (so no higher education) with home values was initially surprising, but does make sense.

Next we'll generate a heatmap using wages, resident count, and home values.
```{r heatmap}
#Select the data for the heat map
chm <- data %>%
  select(avg_annual_pay, resident_count, avg_home_value)

#Calculate the correlation values
cormat <- round(cor(chm),2)

melted_cormat <- melt(cormat)

#Create the Heatmap
library(ggplot2)
ggheatmap <- ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()

reorder_cormat <- function(cormat){
# Use correlation between variables as distance
dd <- as.dist((1-cormat)/2)
hc <- hclust(dd)
cormat <-cormat[hc$order, hc$order]
}

ggheatmap + 
geom_text(aes(Var2, Var1, label = value), color = "black", size = 4)
```

The 52% correlation for wages is strong, but not as strong as expected.  What's surprising is that average pay is showing a lower correlation than education in the previous heatmap.

<B> Home Values Compared to Education, Wages, and Resident Count Colored by Census Region </B>   
There are 2,745 counties across 50 states, but only 4 census regions. We'll use regions to create several scatter charts of the relationship between education, wages, resident count, and home values. 
```{r scatter-charts}
scatterdf <- select(data, region, region_num, hs_diploma_percent, four_year_degree_percent, avg_annual_pay, resident_count, avg_home_value)

#Scatter chart of hs diploma percent
hsd_scatter <- ggplot(scatterdf, aes(avg_home_value, hs_diploma_percent, colour = region)) + 
  geom_point(alpha=0.5, position = 'jitter') +
  scale_x_continuous(labels = comma) + 
  scale_y_continuous(labels = comma)

#Scatter chart of four year degree percent
fyd_scatter <- ggplot(scatterdf, aes(avg_home_value, four_year_degree_percent, colour = region)) + 
  geom_point(alpha=0.5, position = 'jitter') +
  scale_x_continuous(labels = comma) + 
  scale_y_continuous(labels = comma)

#Scatter chart of average annual pay
aap_scatter <- ggplot(scatterdf, aes(avg_home_value, avg_annual_pay, colour = region)) + 
  geom_point(alpha=0.5, position = 'jitter') +
  scale_x_continuous(labels = comma) + 
  scale_y_continuous(labels = comma)

#Scatter chart of resident count
rc_scatter <- ggplot(scatterdf, aes(avg_home_value, resident_count, colour = region)) + 
  geom_point(alpha=0.5, position = 'jitter') +
  scale_x_continuous(labels = comma) + 
  scale_y_continuous(labels = comma)

#Show scatter charts
ggarrange(hsd_scatter, fyd_scatter, aap_scatter, rc_scatter, 
                    labels = c("hsd", "fyd", "aap", "rc"),
                    ncol = 2, nrow = 2,
                    common.legend = TRUE, legend = "bottom")
```

We'll conclude this visuals section with comparisons of the violin plots   
<I> Home Values </I>
```{r violin-plots}
#Violin of home values
hv_violin <- ggplot(scatterdf, aes(region_num, avg_home_value, colour = region)) + 
  geom_violin(trim = FALSE) 

#Display the violin plots
ggarrange(hv_violin,
          labels = c("hv"),
          ncol = 1, nrow = 1,
          common.legend = TRUE, legend = "top")
```

<I> Education </I>
```{r}
#Violin of high school diploma
hsd_violin <- ggplot(scatterdf, aes(region_num, hs_diploma_percent, colour = region)) + 
  geom_violin(trim = FALSE)

#Violin of four year degree
fyd_violin <- ggplot(scatterdf, aes(region_num, four_year_degree_percent, colour = region)) +
  geom_violin(trim = FALSE)

#Display the violin plots
ggarrange(hsd_violin, fyd_violin,
          labels = c("hsd", "fyd"),
          ncol = 2, nrow = 1,
          common.legend = TRUE, legend = "bottom")
```

<I> Wages and Resident Count </I>
```{r}
#Violin of wages
aap_violin <- ggplot(scatterdf, aes(region_num, avg_annual_pay, colour = region)) + 
  geom_violin(trim = FALSE) + 
  scale_y_continuous(labels = comma)

#Violin of resident count
rc_violin <- ggplot(scatterdf, aes(region_num, resident_count, colour = region)) + 
  geom_violin(trim = FALSE) + 
  scale_y_continuous(labels = comma)

#Display the violin plots
ggarrange(aap_violin, rc_violin,
          labels = c("aap", "rc"),
          ncol = 2, nrow = 1,
          common.legend = TRUE, legend = "bottom")
```
