---
title: "Data 607 Project 2, File 2"
author: "Ken Popkin"
date: "3/9/2020"
output: html_document
---
## Assignment Summary
For Project 2, choose any three of the “wide” datasets identified in the Week 6 Discussion items. This document contains the third of the three files selected for this assignment.

File selected: Unicef data on Under 5 Mortality Rates submitted by Samuel Bellows.  [link to Unicef Under 5 Mortality Rates](https://bbhosted.cuny.edu/courses/1/SPS01_DATA_607_02_1202_1/db/_36347024_1/unicef-u5mr.csv)

Analysis to perform: This UNICEF dataset gives the under 5 mortality for many countries across the years 1950-2015. The problem is that the year variable is spread out into 65 different columns, 1 for each year, that need to be gathered into 1 column. Produce a 3 column dataset of country, year, and mortality. Provide narrative descriptions of the data cleanup work and analysis completed, along with conclusions about the data.

## Setup and load data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r}
library(dplyr)
library(tidyr)
library(data.table)
library(ggplot2)
```

```{r, Load-text-file}
#Load the data
data <-  read.csv('unicef-u5mr.csv')
head(data,4)
```

## Tidy up the data
```{r}
#Convert the years data from a wide to long format
yearlydata <- pivot_longer(data,c(2:67), 'year')
yearlydata <- yearlydata %>% `colnames<-`(c('country', 'year', 'mortality'))
yearlydata_sorted <- yearlydata[order(yearlydata$mortality),]
```

## Analysis

```{r}
#Generate some summary information about the dataset
summary(yearlydata)

missing_data <- sum(is.na.data.frame(yearlydata))
percent_missing_data <- round(missing_data/length((yearlydata$year)) * 100)

cat('\n')
cat('Number of country-years missing data:', missing_data,'\n')
cat('Percent of missing data:', percent_missing_data, '%\n')

```

```{r}
mortality_avg <- data %>%
             transmute(CountryName,
             Mean = rowMeans(select(., c(2:67)),na.rm = TRUE))

mortality_avg <- mortality_avg[order(mortality_avg$Mean, decreasing = FALSE),]
head(mortality_avg)
```

```{r}
#Boxplot the data
ggplot(mortality_avg, aes(x = "", y = Mean)) + 
  geom_boxplot()
```

```{r}
mortality_avg_top20 <- tail(mortality_avg,20)
mortality_avg_top20
```


```{r}
#Bargraph the top 20 countries with the highest average mortality
ggplot(mortality_avg_top20, aes(x=reorder(CountryName, Mean), y=Mean)) + 
  geom_bar(stat="identity", fill="lightgreen", color="grey50") +
  ggtitle("Countries with Highest Mortality Rates") +
  xlab("country") + ylab("mortality rate") + 
  coord_flip()
```

```{r}
mortality_avg_bottom20 <- head(mortality_avg,20)
mortality_avg_bottom20
```


```{r}
#Bargraph the countries with the lowest mortality rates
ggplot(mortality_avg_bottom20, aes(x=reorder(CountryName, -Mean), y=Mean)) + 
  geom_bar(stat="identity", fill="lightgreen", color="grey50") +
  ggtitle("Countries with Lowest Mortality Rates") +
  xlab("country") + ylab("mortality rate") + 
  coord_flip()
```

```{r}
#What is the US average mortality rate
subset(mortality_avg, CountryName == 'United States of America')
```

## Conclusions and Recommendations
There are 196 countries in the dataset and the average mortality rate for under_5_years_of_age across all countries is 85 with a median of 54 for the years 1950 to 2015.  50% of the countries lie in a range of 20 to 130.

Mali has the highest mortality rate of 273 and the rate drops to 183 for the 20th highest country.  Moving to lowest mortality rates, Andorra has a rate of 5 and the 20th best country is Estonia at 14.  Interestingly, the United States ranks 33rd with a mortality rate of 18. 

Recommendations for further study of this data are to...
1. align each country with its continent and evaluate countries by continent, similar to prior datasets in this assignment when states were analyzed by region.

2. identify additional demographics about each of these countries and look for the top features that could help to predict a countries mortality rate for children under age 5 (high correlations would be the desired find).