---
title: "DATA 607 Week 1 Assignment"
output: html_notebook

introduction: 'How Popular is Donald Trump? To answer this question FiveThirtyEight.com relies on the dataset trump-approval-ratings (link:  https://projects.fivethirtyeight.com/trump-approval-ratings). This dataset tracks the results of a number of known (i.e. Gallup, Rasmussen) and lesser known polls taken since 2017. Information provided includes the poll date, sample size, and approval and disapproval ratings.'

approach: 'In this notebook, several columns are selected; then the Gallup poll data rows are used to create a chart showing how President Trumps approval numbers have varied the past three years.'
---

'Get the data'
```{r}
approval_data <- read.csv('https://raw.githubusercontent.com/krpopkin/DATA607-Class-Repository/master/Week%201/approval_polllist.csv')

head(approval_data,2)
```

'Obtain the desired columns'
```{r}
library(dplyr)
chart_data <- select(approval_data, startdate, pollster, approve)

head(chart_data,2)
```

Gallup runs a poll almost every day.  To make the chart look better we're only going to use polls on the 15th of each month
```{r}
library(lubridate)
chart_data2 <- mutate(chart_data, startdate2 = mdy(startdate))
chart_data2 <- mutate(chart_data2, day_of_month = day(startdate2))
chart_data2$startdate <- NULL 

head(chart_data2,2)
```

Select the Gallup polls
```{r}
chart_data3 = subset(chart_data2, (pollster == 'Gallup' & day_of_month == 15))
head(chart_data3,2)
```

Create chart for past three years
```{r}
library(ggplot2)
ggplot(data=chart_data3, aes(x=startdate2, y=approve,group=1)) +
  geom_line(color="#aa0022", size=1) +
  geom_point() +
  ggtitle("Gallup Polls 2017 - 2019, Mid-month") +
  labs(x="Year", y="Approval")
```

Key Points: This is a view of only one of may polls, but it does show several points of interest:
1. President Trump's approval has ranged from a low of around 34% in late 2018 to a high of around 43% in late 2019.

2. As of Jan 2020, he is at 39%, which is almost exactly where this poll shows him when it first started in Jan 2017.

3. In spite of all the impeachment drama in recent months, he remains around 6% higher than the lowest approval in late 2018.

Future Work:
1. Continue to monitor the approval rating from this poll in the coming months

2. Layer on several other polls to see how closely they align with Gallup (Rasmussen and Quinnipac are good candidates for this)
