---
title: "Movie Surveys"
author: "Ken Popkin"
date: "2/3/2020"
output:
  pdf_document: default
  html_document: default
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
options(warn=-1)
```
This notebook retrieves data from the movie surveys database and uses the information retrieved to answer the following questions:
Surveys:
1. How many surveys have been completed?

Participants in Surveys
2. How many participants were in each survey?
3. What is the min, max, and average age of the participants?
4. What is the distribution of favorite genres among the survey participants?

The movies included in the survey:
5. What is the distribution of responses for each movie?

```{r}
library(RMySQL)
library(dplyr)
library(ggplot2)
```

User and Password information is removed from this markdown.
```{r setup, include=FALSE}
moviesdb = dbConnect(MySQL(), user='removed from this markdown', password='removed from this markdown', dbname='data607_movies', host='localhost')
```

# Create a dataframe of each table in the movies database
```{r}
rs = dbSendQuery(moviesdb, "select * from people")
dfpeople = fetch(rs, n=-1)

rs = dbSendQuery(moviesdb, "select * from questions")
dfquestions = fetch(rs, n=-1)

rs = dbSendQuery(moviesdb, "select * from responses")
dfresponses = fetch(rs, n=-1)

rs = dbSendQuery(moviesdb, "select * from surveys")
dfsurveys = fetch(rs, n=-1)
```

# Surveys:
## 1. How many surveys have been completed?

```{r}
cat(nrow(dfsurveys), 'survey/s completed')
```

# Participants in Surveys
## 2. How many participants were in each survey?
## 3. What is the min, max, and average age of the participants?

```{r}
dfpeople_age <- dfpeople %>% filter(!is.na(person_age))
dfpeople_age <- select(dfpeople_age, person_age)
```

```{r}
cat('The number of participants in the survey was', nrow(dfpeople), '\n')
cat('The participant ages ranged from', min(dfpeople_age[,1]),'to',max(dfpeople_age[,1]),',with an average age of ', mean(dfpeople_age[,1]) )
```

# Genre's
## The preferred genre's of the participants is illustrated in the chart below...
```{r}
dfpeople_genre <- dfpeople %>% filter(!is.na(person_movie_genre))
dfpeople_genre <- select(dfpeople_genre, person_movie_genre)
```

```{r}
qplot(dfpeople_genre$person_movie_genre, stat = 'count', main = 'Histogram for Genre', xlab = "Genre", fill=I("blue"))

```
# Responses from Participants

## 5. What is the distribution of responses for each movie?

```{r}
dfquestions2 <- select(dfquestions, question_id, question_short)
```

```{r}
dfresponses2 <- select(dfresponses, question_id, response)
```

```{r}
dfcombined <- merge(x=dfresponses2, y=dfquestions2, by = 'question_id', all.x = TRUE)
```

```{r}
dfcombined_count <- summarise_at(group_by(dfcombined,question_short,response),vars(question_id),funs(n()))
head(dfcombined_count, 2)
```


```{r}
library(reshape2)

p <- ggplot(dfcombined_count, aes(x=response, y=question_id))

p <- p + geom_bar(stat = "identity", color = 'darkgreen', fill = 'darkgreen', width = 0.2)

# Divide by levels of "sex", in the vertical direction
p <- p + facet_grid(question_short ~ .)

p
```
